Decision Assistance and Steering in Health Insurance

### 2021-06-11 finals/paper.tex
Updates from Overleaf

\documentclass[12pt]{article}
\usepackage{graphicx,amssymb,amsmath,setspace,comment,verbatim,titling,pgf,lscape,color}
\usepackage[left=2cm,right=2cm,top=2.5cm,bottom=2cm]{geometry}
\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{array}
\usepackage{bbm}
\usepackage{marginnote}
\usepackage[justification=centering]{caption}
%%\usepackage{breqn}
\newcommand{\pderiv}[2]{\frac{\partial#1}{\partial#2}}
%\usepackage{siunitx}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}
\hypersetup{colorlinks,%
						citecolor=black,%
						filecolor=black,%
						linkcolor=black,%
						urlcolor=blue,%
						}
\setstretch{1.5}

@@

\begin{document}
\title{Decision Assistance and Steering in Health Insurance}
\author{Ian McCarthy \& Evan Saltzman}
\date{May 2021}
\maketitle

@@

Our final data consist of 8.3 million total annual enrollments from 2014-2019. Of these, 4.5 million are newly enrolled (e.g., not enrolled in any California exchange plan in the prior year). Table~\ref{tab:summary-stats} displays summary statistics for Covered California households, separately between those that used and did not use decision assistance. The silver tier is the most commonly selected option because consumers eligible for cost sharing reductions (CSRs) must choose a silver plan to receive CSRs.  Approximately 68 percent of California enrollees and 61 percent of Washington enrollees are eligible for CSRs, while 91 percent of California enrollees and 85 percent of Washington enrollees are eligible for premium subsidies.  The proportion of consumers exempt from the individual mandate is small, but is notably higher among those who are uninsured.  The uninsured rate is substantially higher among young adults and males in both states. Individuals with incomes between 250 and 400 percent of FPL, who receive relatively small subsidies, and those with incomes above 400 percent of FPL, who are ineligible for subsidies, make up a relatively large share of the uninsured population.

\begin{table}[h]
\centering	
\caption{Summary Statistics for Covered California Enrollees}
\input{tables/summary-stats.tex}
\footnote{Something}
\label{tab:summary-stats}
\end{table}

@@

For our nested logit discrete choice model, we assume that household $i$ chooses the plan $j$ that maximizes their utility,
\begin{equation}
u_{ij} = \alpha_{i}p_{ij} + \beta x_{j} + \gamma d_{i} + \xi_{j} + \varepsilon_{ij}, \label{eq:utility}
\end{equation}
where $p_{ij}$ denotes the premium for plan $j$, $x_{j}$ denotes a vector of product characteristics, $d_{i}$ denotes a vector of household characteristics, $\xi_{j}$ captures unobserved product characteristics, and $\varepsilon_{ij}$ is an error term assumed to follow a type I extreme value distribution. We allow for heterogenities in price elasticities across households by parameterizing $\alpha_{i}$ in Equation \eqref{eq:utility}, $\alpha_{i} = \alpha + \theta d_{i}.$ In our estimation, this amounts to including a full set of interactions between $d_{i}$ and $p_{ij}$.

this estimator proceeds in four steps: 1) estimate a logit regression of decision assistance, the predicted probabilities of which enter as weights in the second stage; 2) estimate a weighted nested logit model of plan choice, limiting the sample to those observed to \textbf{not} have used any form of decision assistance; 3) use the results of our choice model to predict the counterfactual outcome of plan choice among those that \textbf{did} use some form of decision assistance; and 4) take the difference in predicted plan choices versus observed plan choices, conditional on receiving assistance.

@@

Interpreting our results as an estimate of the ATT assumes that individuals receiving decision assistance do not seek assistance due to some unobserved factors that also affect their plan choice. Conditional on household characteristics as well as insurer fixed effects, interacted with a full set of region and year fixed effects, we argue that this assumption is plausible.

\section{Dominated Choices}
\label{sec:dominated}
As a complementary empirical analysis, we identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set, and we estimate the effects of decision assistance on the probability of making such a choice. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year, insurer, and region fixed effects.

@@

\subsection{Model of Commissions and Steering}
\label{subsec:steering-model}

\subsection{Structural Estimation}
\label{subsec:steering-methods}

>>>>

\documentclass[12pt]{article}
\usepackage{graphicx,amssymb,amsmath,setspace,comment,verbatim,titling,pgf,lscape,color}
\usepackage[left=2cm,right=2cm,top=2.5cm,bottom=2cm]{geometry}
\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{array}
\usepackage{bbm}
\usepackage{marginnote}
\usepackage[justification=centering]{caption}
%%\usepackage{breqn}
\usepackage{booktabs}
\newcommand{\pderiv}[2]{\frac{\partial#1}{\partial#2}}
%\usepackage{siunitx}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}
\hypersetup{colorlinks,%
						citecolor=black,%
						filecolor=black,%
						linkcolor=black,%
						urlcolor=blue,%
						}
\setstretch{1.5}

@@

\begin{document}
\title{Decision Assistance and Steering in Health Insurance}
\author{Ian McCarthy \& Evan Saltzman}
\date{June 2021}
\maketitle

@@

Our final data consist of 8.3 million total households from 2014-2019. Of these, 4.5 million were not previously enrolled in any Covered California plan in the prior year. Table~\ref{tab:summary-stats} displays summary statistics for all households in our data. Since the key independent variable throughout much of our analysis is whether a household used some form of decision assistance, we present summary statistics separately between those that used and did not use decision assistance.

We see from Table~\ref{tab:summary-stats} that enrollees using decision assistance are of lower income and larger household size. Among types of plans purchased, we see little difference between those with and without decision assistance, with the exception of Kaiser plans for which individuals using decision assistance are less likely to purchase. Overall, the silver tier is the most commonly selected option, and this is particularly true for those using decision assistance. Enrollees with decision assistance are also less likely to select an HMO and relatively more likely to select a PPO. Finally, among those using some form of decision assistance, about 70\% of households used a private agent or broker while the remaining 30\% used a public navigator.

@@

For our nested logit discrete choice model, we assume that household $i$ chooses the plan $j$ that maximizes their utility,
\begin{equation}
u_{ij} = \alpha_{i}p_{ij} + \beta x_{j} + \gamma d_{i} + \xi_{j} + \varepsilon_{ij}, 
\label{eqn:utility}
\end{equation}
where $p_{ij}$ denotes the premium for plan $j$, $x_{j}$ denotes a vector of product characteristics, $d_{i}$ denotes a vector of household characteristics, $\xi_{j}$ captures unobserved product characteristics, and $\varepsilon_{ij}$ is an error term assumed to follow a type I extreme value distribution. We allow for heterogenities in price elasticities across households by parameterizing $\alpha_{i}$ in Equation~\eqref{eqn:utility}, $\alpha_{i} = \alpha + \theta d_{i}.$ In our estimation, this amounts to including interactions between $d_{i}$ and $p_{ij}$.

Using the nested logit model in Equation~\eqref{eqn:utility}, our goal is to estimate the average treatment effect on the treated (ATT), which in our case reflects the average treatment effect among those using some form of decision assistance. To allow for general heterogeneous treatment effects, we estimate the ATT with a four-step process: 
\begin{enumerate}
    \item Estimate the nested logit model of plan choice from Equation~\eqref{eqn:utility}, limiting the sample to those observed to \textbf{not} have used any form of decision assistance. We estimate Equation~\eqref{eqn:utility} separately for each year and region, implicitly allowing for a full set of region/year interactions among all covariates. For this reason, we select a relatively small set of covariates, including household-specific premiums, the interaction between premiums and household size, the actuarial value of the plan, the metal tier, indicators for network type (HMO or PPO), and insurer fixed effects. We then obtain a set of coefficient estimates, $\hat{\theta} = \{\hat{\alpha}, \hat{\beta}, \hat{\gamma}, \hat{\xi} \}$.\footnote{These coefficient estimates differ by region and year, but we supress that notation for brevity.}

    \item Use $\hat{\theta}$ to form the predicted counterfactual, $\widehat{Pr[Y_{ij}^{0} | \text{Assistance}]}$. In other words, we use the results of our choice model, estimated only among those without decision assistance, to predict the counterfactual outcome of plan choice among those that \textbf{did} use some form of decision assistance.

    \item Sum the predicted probabilities across individuals, $$\widehat{N_{j}^{0}} = \sum_{ i \in \Omega} \widehat{Pr[Y_{ij}^{0} | \text{Assistance}]},$$ where $\Omega$ denotes the set of all enrollees that used decision assistance. Among such enrollees, this reflects the predicted number of enrollments if enrollees had not used any assistance.

    \item Sum the observed number of enrollments, $N_{j}^{1}$, and take the difference between this and the counterfactual prediction,
    \begin{equation}
        ATT_{j} = \underbrace{N^{1}_{j}}_{Observed} - \underbrace{\widehat{N_{j}^{0}}}_{Predicted}.
    \label{eqn:att}
    \end{equation}
\end{enumerate}

Interpreting Equation~\eqref{eqn:att} as an estimate of the ATT assumes that individuals receiving decision assistance do not seek assistance due to some unobserved factors that also affect their plan choice. Conditional on household characteristics as well as insurer fixed effects, (implicitly) interacted with a full set of region and year fixed effects, we argue that this assumption is plausible; however, our estimation strategy employs untreated households (i.e., those without decision assistance) to predict counterfactual choices for treated households and therefore requires sufficient balance between treated and untreated households. As is evident from Table~\ref{tab:summary-stats}, there is reason to suspect that such balance does not hold in our data.

To ensure balance among observed covariates, we employ an augmented inverse propensity weighted regression estimate \cite{wooldridge2010}, in which the weights in the nested logit estimation derive from a preliminary logit regression of decision assistance status for each household. The outcome in this preliminary regression is an indicator variable, set to one if a household used decision assistance (i.e., the treatment group) and 0 otherwise (i.e., the control group). Denoting the resulting predicted probability of decision assistance by $\hat{p}_{i}$, we then assign a weight of $\frac{\hat{p}_{i}}{1-\hat{p}_{i}}$ to each observation in the control group and proceed with step (1) above using these weights in the nested logit regression. Conceptually, this approach assigns more weight to households without decision assistance that are ``most similar'' to households that use decision assistance.

@@

Figure~\ref{fig:ps-balance} presents the resulting propensity scores from our preliminary logit regression of decision assistance against observable household demographics. As is evident from the figure, we observe substantial overlap in propensity scores between the treated and control groups. In Figure~\ref{fig:cov-balance}, we present a comparison of observable household characteristics between the treatment and control groups. The hollow circles in Figure~\ref{fig:cov-balance} reflect the unadjusted differences in means, while the solid circles reflect the difference in means when weighting households. Weights for treated households are set to 1, and weights for untreated households are set to  $\frac{\hat{p}_{i}}{1-\hat{p}_{i}}$ based on the preliminary logistic regression.  We see from Figure~\ref{fig:cov-balance} that our treated and control groups are well-balanced after weighting.

\section{Dominated Choices}
\label{sec:dominated}
The results in Section~\ref{subsec:causal-results} reveal large effects of decision assistance on plan choice; however, it is not clear from those results whether decision assistance \textit{improves} plan choice or simply reallocates households into different, perhaps worse, plans.

we identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set, and we estimate the effects of decision assistance on the probability of making such a choice. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year, insurer, and region fixed effects.

@@

\subsection{Model}
\label{subsec:steering-model}

Consider a two-stage  game where  (1) firms set premiums and commissions simultaneously to maximize their expected profit and (2) consumers select a plan to maximize their expected utility. In the following three subsections, we detail how we (1) model the decisions of firms; (2) model the decisions of consumers; and (3) find the model equilibrium.

Given our available data, we do not explicitly model the actions of individual insurance agents.  However, we do model how firms use agent compensation to steer an agent's clients to their plans.  Our approach assumes that the consumer's decision to seek agent assistance is independent of the firms' commission levels.  This identifying assumption is valid if commissions incentivize agents to recommend certain plans to their clients, but do not induce agents to change advertising levels or marketing efforts to  seek new clients. Advertising and marketing by both insurers and agents are tightly regulated, particularly in Covered California.  We therefore believe our identifying assumption is relatively  innocuous.

\subsubsection{Firms Set Premiums and Commissions}

\paragraph{Defining Firm Decision Variables}

Firms set premiums and commissions to maximize their expected profit.  Let $p_{jmt}$ be the base premium decision for each plan $j$ sold in market $m$ in period $t$ and the \textit{commission decision} $n_{gjt}$

	\vspace{-0.4in}		
		\singlespacing

				\begin{eqnarray*}
	n_{gjt}  \equiv \begin{cases} 
		\delta_{gjt} & \mathrm{\% \,\, commission} \\		
		\eta_{gjt} & \mathrm{\$ \,\, commission} \\
		0 & \mathrm{o/w}
   \end{cases}
	\end{eqnarray*}

\doublespacing

\noindent for each plan $j$ sold at time $t$ with renewal status $g$ (i.e., new or existing customer). Agents are either paid a percentage $\delta_{gjt}$  of the consumer's unsubsidized premium  or a flat amount $\eta_{gjt}$ that is independent of the premium.  We assume the firm's decision to pay a percentage commission or flat commission is exogenous because switching agent compensation methods is costly.  The only firm that switched between compensation methods during our six-year study timeframe was Health Net following its acquisition by Centene. The \textit{commission paid} to household $i$'s insurance agent is

		\vspace{-0.4in}		
		\singlespacing

			\begin{eqnarray*}
	\eta_{gijt}(\textit{p},\textit{n})  \equiv \begin{cases} 
		n_{gjt} \left(\sigma_{it}p_{jmt}\right) & \mathrm{\% \,\, commission} \\		
		n_{gjt} & \mathrm{\$ \,\, commission}
   \end{cases}
	\end{eqnarray*}

\doublespacing

\noindent where $\textit{p} = {\bf p}_t$ is the vector of plan base premiums set by all insurers in each market in year $t$, $\textit{n} = {\bf n}_t$ is the vector of commission decision variables in period $t$, and the household's (unsubsidized) premium is the product of the household's rating factor $\sigma_{it}$  and the plan's base premium $p_{jmt}$. In California, the rating factor can vary by age and geographic residence of the household's members.\footnote{The ACA also permits rating by tobacco usage, but California prohibits tobacco rating.}  Insurers can charge a 64-year-old up to 3 times as much as a 21-year-old. Figure~\ref{rating_area_partitions} shows the partition of California into rating areas.  In each rating area,  an insurer's premium must be the same for all consumers of the same age. Total commissions paid by firm $f$  are

	\vspace{-0.2in}

		$$
		\mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) = \sum_{g \in G, i \in I, m\in M, k \in J_{fmt}} \left( \mathbb{I}_{n,i,t} \right)  q_{ikt}(\textit{p},\textit{n};\boldsymbol{\beta}) \eta_{gikt}(\textit{p},\textit{n})
		$$

\noindent where $\mathbb{I}_{n,i,t}$ indicates whether household $i$ is assisted by an agent in year $t$, $q_{ikt}(\cdot)$ is the probability household $i$ chooses plan $k$ at time $t$, and $J_{fmt}$ is the set of plans offered by firm $f$ in market $m$ at time $t$.  The model parameters $\boldsymbol{\theta} \equiv (\boldsymbol{\beta},\boldsymbol{\gamma},\boldsymbol{\mu})$, where $\boldsymbol{\beta}$ are the demand parameters, $\boldsymbol{\gamma}$ are the risk score parameters, and $\boldsymbol{\mu}$ are the average claims parameters.

\begin{figure}[htbp]
\caption{Premium Rating Regions in California}
\label{rating_area_partitions}
\centering
\includegraphics[scale=.45]{finals/pics/CA_Rating_Regions2.jpg}

\begin{footnotesize}
 \begin{flushleft}
Notes: Figure shows the premium rating regions in the California exchange \citep{CARATE2016}. California's 58 counties are divided into 19 rating areas. 
\end{flushleft}
\end{footnotesize}

\end{figure}

\paragraph{Calculating Risk Adjustment Transfers}

We now discuss the ACA's risk adjustment program, which directly affects firm profit. \citet{Pope2014} derive the ACA risk adjustment transfer formula.\footnote{We start with \citet{Pope2014}'s transfer formula as derived in their first appendix, which allows plans to vary only by their actuarial values (and not by differences in firm efficiency, geographic costs, allowable rating factors, or moral hazard).  We start with this formula because we want to capture  all differences in expected risk, except for cost sharing and any associated moral hazard, in the plan's risk score (i.e., cost sharing and moral hazard are addressed through the utilization share).  In contrast, the plan liability risk score   $PLRS_j$ as defined in \citet{Pope2014}'s second appendix does not account for certain differences such as variation in geographic cost.  Instead, \citet{Pope2014} account for these differences by applying factors in the transfer formula.}  In our notation, the per-member per-month risk adjustment transfer $ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$ for plan $j$ is

\vspace{-0.2in}

	$$
	ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) = \left(\frac{r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})\sum_{m \in M, l \in J_{mt}} q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}    -     \frac{h_j\sum_{m \in M, l \in J_{mt}} q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} h_l q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}\right)\overline{p}
	$$

\noindent where $r_{jmt}(\cdot)$ is the plan risk score, $h_j$ is an exogenous expected utilization factor set by regulation that accounts for the plan AV and any associated moral hazard, and $\overline{p}$ is the weighted average premium in the market.  The plan's total risk adjustment transfer $RA_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$ equals

	\vspace{-0.4in}

	\begin{eqnarray}\label{ra_formula}
	RA_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta}) \nonumber \\
	&= & \left[rs_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  - us_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  \right] R_t(\textit{p},\textit{n};\boldsymbol{\beta}) 
	\end{eqnarray}

	\vspace{-0.2in}

\noindent where $rs_{jmt}(\cdot)$ is the plan's ``risk share'' of total claims, $us_{jmt}(\cdot)$ is the plan's ``utilization share'' of total claims, and $R_t(\cdot) = \sum_f R_{ft}(\cdot)$ is total premium revenue across all plans. The total risk adjustment transfer for the firm equals the sum of the risk adjustment transfers for the plans that it sells (i.e., $RA_{ft}(\cdot) = \sum_{m \in M, j \in J_{fmt}} RA_{jmt}(\cdot)$.

The plan's risk share includes the combined effects of adverse selection, moral hazard, and plan AV, whereas the plan's utilization share only includes the effects of moral hazard and plan AV. Thus, the difference between the risk share and utilization share captures the plan's relative risk due to adverse selection only.   The risk share equals

\vspace{-0.1in}

	$$
	rs_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) = \frac{ r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}}  r_{lmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}
	$$

\noindent where $J_{mt}$ is the set of all plans offered in market $m$ and year $t$, and the plan risk score $r_{jmt}(\cdot)$ is a function of enrollee characteristics and the plan AV. We do not directly observe plan risk scores in the insurer rate filing data.  However, we observe all other variables in formula~(\ref{ra_formula}) including each plan's risk adjustment transfer and can therefore solve formula~(\ref{ra_formula}) for the plan risk scores. The plan's utilization share $us_{jmt}(\cdot)$ equals

\vspace{-0.1in}

	$$
	us_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) =  \frac{ h_j q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} h_l q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}
	$$

\noindent A positive difference between the risk share and the utilization share indicates that a plan has high risk relative to its expected utilization and results in the plan receiving a risk adjustment transfer; a negative difference indicates that a plan has low risk relative to its expected utilization and results in the plan paying a risk adjustment transfer.

\paragraph{Calculating Variable Profit}

 Firms choose premiums and commissions to maximize expected variable profit

	\vspace{-0.4in}		
	\begin{footnotesize}

	\begin{eqnarray}\label{profit}
	\pi_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) &=&  R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) -  C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) + RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) +  RI_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})  - V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})  \nonumber \\
	&=& R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - (1 - \iota_{ft}) C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) + RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})   - V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) 
	\end{eqnarray}

	\end{footnotesize}

	\vspace{-0.1in}

\noindent where $R_{ft}(\cdot)$ is total premium revenue, $C_{ft}(\cdot)$ is total claims, $RI_{ft}(\cdot)$ is reinsurance received, $V_{ft}(\cdot)$  is variable administrative cost (excluding commissions), and $\iota_{ft}$ is the actuarial value of the reinsurance contract (i.e., the expected percentage of claims paid by the reinsurer).  Total commissions paid $\mathcal{C}_{ft}(\cdot)$ and risk adjustment received  $RA_{ft}(\cdot)$ are as above.

\subsubsection{Households Choose Plans}

Households choose the plan that maximizes their (indirect) utility function

	\vspace{-0.4in}

	\begin{eqnarray}\label{demand_model}
	U_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) &\equiv& \beta_{i}^p p_{ijt}(\textit{p}) + \beta_j^{\eta} \eta_{gijt}(\textit{p},\textit{n})   + \beta_i^y y_{ij(t-1)} + x'_{ij}\beta^x + \xi_{j} + \epsilon_{ijt}^d
	\end{eqnarray}

	\vspace{-0.1in}

\noindent where $\textit{p} = {\bf p}_t$ is the vector of plan base premiums set by all insurers in each market in year $t$, $p_{ijt}(\textit{p})$ is household $i$'s premium for plan $j$ in year $t$, $y_{ij(t-1)}$ indicates whether household $i$ chose plan $j$ in the previous year, $x_{ij}$ is a vector of observed product characteristics including the plan AV,  $\xi_{j}$ is a vector of unobserved product characteristics, and $\epsilon_{ijt}^d$ is an error term.  We allow the household's premium parameter $\beta_{i}^p = \beta^p + w'_{it} \phi$ to vary with household characteristics $w_{it}$ and the household's inertia parameter $\beta_i^y = \beta^y +  x'_{ij} \kappa + w'_{it} \nu$ to vary with household and product characteristics. CSRs enter utility equation (\ref{demand_model}) through the plan AV and premium subsidies reduce the household's premium $p_{ijt}(\textit{p})$.   The utility of the outside option $U_{i0t} = \beta_i^p \rho_{it} + \epsilon_{i0t}$, where $\rho_{it}$ is the household's penalty for not having insurance in year $t$.

\paragraph{Calculating Household Premiums}

Households choose the plan that maximizes their (indirect) utility function

	\vspace{-0.4in}

	\begin{eqnarray}\label{demand_model}
	U_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) &\equiv& \beta_{i}^p p_{ijt}(\textit{p}) + \beta_j^{\eta} \eta_{gijt}(\textit{p},\textit{n})  \mathbb{I}_{n,i,t}   + \beta_i^y y_{ij(t-1)} + x'_{ij}\beta^x + \xi_{j} + \epsilon_{ijt}^d
	\end{eqnarray}

	\vspace{-0.1in}

\noindent where $p_{ijt}(\textit{p})$ is household $i$'s (subsidized) premium for plan $j$ in year $t$, $\eta_{gijt}(\textit{p},\textit{n})$ is the commission paid to household $i$'s insurance agent, $y_{ij(t-1)}$ indicates whether household $i$ chose plan $j$ in the previous year, $x_{ij}$ is a vector of observed product characteristics including the plan AV,  $\xi_{j}$ is a vector of unobserved product characteristics, and $\epsilon_{ijt}^d$ is an error term.  CSRs enter utility equation (\ref{demand_model}) through the plan AV and premium subsidies reduce the household's premium $p_{ijt}(\textit{p})$.   The utility of the outside option $U_{i0t} = \beta_i^p \rho_{it} + \epsilon_{i0t}$, where $\rho_{it}$ is the household's penalty for not having insurance in year $t$.

\paragraph{Choice Model Parameters}

We define the vector of demand parameters $\boldsymbol{\beta} \equiv (\beta_i^p,\beta_j^{\eta},\beta_i^y,\beta^x)$ to accommodate heterogeneous consumer preferences.  The household's premium parameter $\beta_{i}^p \equiv \beta^p + w'_{it} \phi$ varies with household characteristics $w_{it}$.  The household's inertia parameter $\beta_i^y = \beta^y +  x'_{ij} \kappa + w'_{it} \nu$ varies with both household and product characteristics. The steering parameter $\beta_j^{\eta}$ captures the degree to which agent compensation steers  an agent's clients to particular plans.  A high value of $\beta_j^{\eta}$ would indicate that agent compensation is effective in steering consumers.   Because smaller firms offer agents considerably higher commissions, we allow the steering parameter to vary with firm size.

\paragraph{Calculating Demand}

 We assume equation (\ref{demand_model}) is a nested logit model such that the vector of error terms $\epsilon_i$ has the generalized extreme value distribution.  The first nest contains all exchange plans and  the second nest contains only the outside option.  This two-nest structure captures the key observed substitution pattern between the silver tier and the outside option resulting from the ACA's linkage of CSRs to the purchase of silver plans.  Under the assumption that $\epsilon_i$ has the generalized extreme value distribution, the household choice probabilities are

	\vspace{-0.4in}

	\begin{eqnarray}\label{choice_probs}
	q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) &=&  \frac{e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\left(\sum_{j } e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\right)^{\lambda-1}}{1 + \left(\sum_{j} e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\right)^{\lambda}} 
	\end{eqnarray}

	\vspace{-0.1in}

\noindent where $V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) \equiv \beta_{i}^p p_{ijt}(\textit{p}) + \beta_j^{\eta} \eta_{gijt}(\textit{p},\textit{n})   + \beta_i^y y_{ij(t-1)} + x'_{ij}\beta^x + \xi_{j}$ and $\lambda$ is the nesting parameter. The household choice probabilities in equation~(\ref{choice_probs}) converge to the standard logit choice probabilities when $\lambda \rightarrow 1$.

The sensitivity of a subsidized consumer's demand $q_{ijt}(\textit{p})$ to a premium change is given by

	\vspace{-0.4in}

	\begin{eqnarray*}
	\frac{\partial q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\partial p_{jmt}} &=& \sum_{l \in J_{mt}} \frac{ \partial q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\partial p_{ilt}(\textit{p})}\frac{ \partial p_{ilt}(\textit{p})}{\partial p_{jmt}}
	\end{eqnarray*}

	\vspace{-0.1in}

\noindent for all plans $j,k$ in the set of available plans $J_{mt}$.  Assuming a strictly positive subsidy that does not exceed the full, unsubsidized premium, it follows from equation (\ref{cons_premium}) that

	\vspace{-0.4in}
	\singlespacing

	\begin{eqnarray}\label{price_partial_formula}
	 \frac{\partial p_{ilt}(\textit{p})}{\partial p_{jmt}} = \begin{cases} 
		0 & l=j, j = b \\		
		\sigma_{it} & l = j, j \neq b \\     
		-\sigma_{it} & l \neq j, j = b \\
		0 & l \neq j, j \neq b

   \end{cases}
	\end{eqnarray}

	\doublespacing

	%\vspace{-0.1in}

\noindent For a non-benchmark plan, an infinitesimal premium increase results in consumers paying more for that plan only.  An infinitesimal increase in the benchmark premium does not affect what subsidized consumers pay for the benchmark plan, but reduces what consumers pay for all other plans because of the larger subsidy. The complex relationship between insurer and consumer premiums, endogenous determination of the benchmark premium, and variation in the benchmark plan across consumers due to heterogeneous entry create significant computational challenges for estimation.  We carefully model the endogenous subsidy design despite the high computational cost because of the critical role premium subsidies play in addressing adverse selection.

\subsubsection{Equilibrium}

Now we find the first-order necessary conditions for a Nash equilibrium.  Differentiating equation (\ref{profit}) with respect to the premium and the commission yields the  first-order conditions

\vspace{-0.4in}	
\begin{footnotesize}

	\begin{eqnarray}\label{acara_foc_short}
	\frac{\partial R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} &=& (1-\iota_{ft}) \frac{\partial C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} \nonumber \\
	\frac{\partial R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} &=& (1-\iota_{ft}) \frac{\partial C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}}
	\end{eqnarray}

\end{footnotesize}

	\vspace{-0.1in}

\noindent for all renewal groups $g$ and markets $m$ in which plan $j$ is offered by the firm in year $t$. In Appendix~\ref{app::model_formulas}, we write every variable in equations~(\ref{profit}) and (\ref{acara_foc_short})  in terms of three estimable variables: (1) household choice probabilities $q_{ijt}(\textit{p},\textit{n};\boldsymbol{\theta})$; (2) plan risk scores $r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$; and (3) average claims $c_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$.  Household choice probabilities are computed using equation (\ref{choice_probs}).  We calculate plan risk scores as a function of observable enrollee characteristics   and the plan generosity using the estimating equation

\vspace{-0.4in}

	\begin{eqnarray}\label{risk_score_estimation}
	\ln r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& \sum_{d \in D} \gamma^d s_{djmt}(\textit{p},\textit{n};\boldsymbol{\theta}) + MT'_j \gamma^{MT} + \epsilon_{jmt}^r
	\end{eqnarray}

\vspace{-0.1in}

\noindent The predicted demographic share $s_{djmt}(\cdot)$ is the share of plan $j$'s enrollment in market $m$ and year $t$ with demographic characteristic $d$, $MT_j$ is a vector metal tier fixed effects,  $\epsilon_{jmt}^r$ is an error term, and the vector of risk score parameters $\boldsymbol{\gamma} = (\gamma^d,\gamma^{MT},\gamma^{n})$. The demographic shares are computed by aggregating the household choice probabilities.   We calculate plan average claims as a function of the plan risk score using the estimating equation

\vspace{-0.4in}

	\begin{eqnarray}\label{avg_claims_estimation}
	\ln c_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& \mu^r \ln r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) + x'_j  \mu^{x}  + \mu^l l_t +   n'_m \mu^{n} + \epsilon_{jmt}^c
	\end{eqnarray}

\vspace{-0.1in}

\noindent where $r_{jmt}(\cdot)$ is the predicted risk score computed using equation (\ref{risk_score_estimation}), $x_j$ are product characteristics (not including plan AV), $l_t$ is a linear trend, $n'_m$ are market fixed effects, $\epsilon_{jmt}^c$ is an error term, and  $\boldsymbol{\mu} = (\mu^r,\mu^x,\mu^l,\mu^n)$ are the claims parameters.

\subsection{Estimation}\label{subsec:steering-methods}

In this section, we explain how we use the generalized method of moments (GMM) to estimate the parameter vector $\boldsymbol{\theta}$.  We construct four sets of moment conditions: (1) demand moments that match observed choices and predicted household choice probabilities; (2) risk score moments that match observed and predicted plan risk scores; and (3) average claims moments that match observed and predicted plan average claims; and (4) the first-order conditions for profit maximization in equation~\ref{acara_foc_short}.  These moment conditions are summarized below:

	\vspace{-0.4in}
